\section{Data Location}\label{data-location}

We talked about value types and reference types. Reference types, which
consist of structs, arrays and mappings in \texttt{Solidity} allow for a
specification of their data location. This is an additional annotation
and it indicates where that reference type variable is stored.

There are three locations: \textbf{memory}, \textbf{storage} and
\textbf{calle data}. Remember that these are 3 of the 4 locations that
the EVM supports besides the stack. These data location affects the
lifetime or the scope and persistence of the variables stored in those
locations.

\begin{itemize}
\item
  \textbf{Memory} indicates that the lifetime is limited to that
  external function call.
\item
  \textbf{Storage} indicates that the lifetime extends to that whole
  contract and this is also the location where state variables are
  stored.
\item
  \textbf{Call data} is a non-modifiable and non-persistent area where
  function arguments are stored. This is required for parameters of
  external functions but can also be used for other variables.\\

  This data location annotation impacts the scope of the variables that
  use this lotation. From a security perspective this affects the
  persistence of those variables.
\end{itemize}

\subsection{Assignments}\label{assignments}

\textbf{The data location annotation} we just talked about not only
\textbf{affects the persistency of those variables}, the scope in which
they are relevant, but it also affects what are known as
\textbf{assignment semantics}.

In the context of \texttt{Solidity}, what this means is that during an
assignment, using such variables is a copy of that variable being
created? Or is simply a reference being created to the existing
variable?

In \texttt{Solidity}, storage to memory assignments always create an
independent copy. Memory to memory assignments only create references.
Similarly storage to storage assignments only create a reference. All
other variants, create a copy.

From a security perspective how this impacts the semantics is: if a copy
were to be created because of these assignment rules, then any
modifications to the copy affect only the copy and not the original
variable from where it was copied.

On the other hand, if a reference was created, in the case of memory to
memory assignments or storage to storage assignments, then the new
variable modifications to that affect the original variable because both
of them are just different names pointing to the same underlying
variable data (the same memory address on the machine).

So this becomes important when you analyze programs and notice what the
data locations are for those reference types, because there's a big
difference if modifications are being made to the copy versus a
reference.
