\section{Shadowing and
Pre-declaration}\label{shadowing-and-pre-declaration}

\subsection{Shadowing}\label{shadowing}

Shadowing of built-in \texttt{Solidity} variables was a concern in some
of the older \texttt{Solidity} versions. Built-in variables such as
\texttt{now}, \texttt{assert} and some others could be shadowed by other
variables, functions or modifiers in the contract to override their
behavior.

This as you can imagine is dangerous and could lead to many unexpected
behavior and therefore this Shadowing was disallowed in later
\texttt{Solidity} versions.

Similar to the Shadowing of built-in variables the older versions of
\texttt{Solidity} also allowed state variable shadowing.

This meant that the right contracts could have state variables that had
the same name as some of their base contracts. You can imagine that the
base variables and shadowed variables with the same names could be
confusing even for the developer and they could end up using or
modifying the wrong variable from the base contracts.

This dangerous and unexpected consequences was recognized, so
\texttt{Solidity} compiler \texttt{0.6.0} disallowed Shadowing of state
variables.

\subsection{Pre-declaration}\label{pre-declaration}

Earlier versions of \texttt{Solidity} allowed the use of local variables
even before they were declared.

These variables could be declared later or they could have been declared
in another scope. This led to undefined behavior as you may expect.

\texttt{Solidity} version \texttt{0.5.0} and beyond change this, to
implement the popular C99-style scoping rules where variables can only
be used after they have been declared and only in the same or nested
scopes.
