\section{Visibility}\label{visibility}

Remember that functions in \texttt{Solidity} have the notion of
visibility where they could be either \texttt{public},
\texttt{external}, \texttt{internal} or \texttt{private}, this affects
which users can call these functions.

So \texttt{public} and \texttt{external} functions are callable by
anyone depending on the access control that is enforced on top of that,
whereas \texttt{internal} and \texttt{private} can be called only from
within the contracts or the derived contracts.

Until \texttt{Solidity} version \texttt{0.5.0} this visibility specifier
was optional and they defaulted to \texttt{public}. This aspect led to
vulnerabilities where the developer forgot to mention or specify the
visibility in which case it became public by default and resulted in
malicious users being able to call these functions and make unauthorized
state changes completely unexpected by the developer or the smart
partner.

So this optional specification of function visibility defaulting to
\texttt{public} visibility was removed as of \texttt{Solidity} version
\texttt{0.5.2}, so this was a big change when it came to increasing the
security of smart contracts and since that version function visibility
is required to be specified explicitly for every function.

\subsection{Public Functions}\label{public-functions}

Remember that \texttt{Solidity} has the notion of visibility for
functions, there are four visibility specifiers: \texttt{internal},
\texttt{private}, \texttt{public} and \texttt{external}. \texttt{public}
functions consume more Gas than \texttt{external} functions.

The reason for this is because the arguments of \texttt{public}
functions need to be copied from thecall data component of the EVM to
the memory component. This copying produces more bytecode for such
\texttt{public} functions which therefore consumes more Gas.

This copying is not required for \texttt{external} functions where their
arguments can be left behind in the calldata component of the EVM. This
key difference leads to \texttt{public} functions consuming more Gas
than \texttt{external} functions in \texttt{Solidity}.

So if there are functions in the contract that are never called from
within the contracts themselves, then such functions should be declared
with \texttt{external} visibility and not \texttt{public} visibility,
which leads to better Gas efficiency.
