\section{Testing, Unused and Redundant
Code}\label{testing-unused-and-redundant-code}

\subsection{Testing}\label{testing}

Software testing or validation is a fundamental software engineering
practice that is a critical contributor to improved security. Testing
validates whether the system implementation meets the requirements as
detailed by the specification. Unit tests, functional tests, integration
and end-to-end tests should have been performed to achieve good test
coverage across the entire code base.

\begin{itemize}
\tightlist
\item
  Changes introduced with any revisions should be validated with
  regression tests.
\item
  Smoke testing indicates at a high level if the functionality works or
  not.
\item
  Stress testing validates extreme scenarios with borderline cases to
  check if those have been considered correctly.
\item
  Performance and security specific testing validates those aspects
  respectively.
\end{itemize}

Any code or parameterization used specifically for testing should be
removed from production code, which in smart contracts may apply
differently to testnets vs.~mainnet. Leaving test parameters or
configurations behind may accidentally allow their usage resulting in
unexpected maintenance behavior or serious vulnerabilities, so overall
we need to ensure that sufficient levels of testing have been performed
across all these different categories we just mentioned.

\subsection{Unused Code}\label{unused-code}

Unused constructs may negatively impact security. This applies to any
unused reports, inherited contracts, functions, parameters, variables,
modifiers, events or return values; all of which should be removed or
used appropriately after careful evaluation.

Removing will not only reduce Gas costs, but also improve readability
and maintainability of the code. Unused constructs may also be
indicative of missing logic that may be a security concern, if that
logic were to have implemented security related functionality, so one
needs to either remove or use such unused constructs.

\subsection{Redundant Code}\label{redundant-code}

Redundant constructs are also concerned. These are a kind of constructs
that are not required either because there are equivalent constructs
that implement the same functionality or because they are not relevant
anymore. Such redundant code and comments can be confusing and should be
removed or changed appropriately after careful evaluation.

Similar to unused constructs, removing redundant constructs will not
only reduce Gas costs but also improve readability and maintainability
of the code. If redundant constructs are indicative of missing or
incorrect logic, then they may be a security concern, if such logic were
to have implemented security related functionality. So one needs to
either remove such redundant constructs or make them relevant by adding
or changing the corresponding logic.
